Paper_Results

### 2025-06-10 Paper_Results.Rmd
Review Changes

## Pitfall 1: Within person performance

```{r }
############ Table 3 #############
##################################

# Basic Parameters from Study
n_features <- 10  # Features
n_samples <- 90  # Timepoints
n_subjects <- 150 # Subjects

# Generating Features
A <- 0.05 # Relationship between features and outcome
feature_std <- 0.1 # population level feature generating process
B <- 0.8  # Cross-subject variability ("random effect") (added per participants for all timepoints)
C <- 0.2 # Within-subject variability (added within participant for each timepoint)

@@

set.seed(12361488)
features_sample <- create_data(n_features,n_samples,n_subjects,A,feature_std,B,C,overall_prob_outcome,sd_outcome,time_effect = FALSE)
features_sample <- features_sample[[1]]

@@

@@

result <- read.csv("simulation_results.csv") # see Run_Simulation.R to investigate how results were created

p1 <- ggplot(result[result$A == 0,], aes(x=icc, y=auc, color = icc_pred)) +
  geom_point(alpha=0.4, size = 0.8) +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#2A363B","#83AF9B","#C8C8A9","#F9CDAD","#FC9D9A","#FE4365")) +
  ylab("Prediction Performance (Overall AUC)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome \n(ICC Outcome)") +
  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
  ggtitle("Bias in Performance Estimates (AUC) When No Relationship Exists: Expected AUC = 0.5") +
  geom_hline(yintercept = 0.5, color = "#83AF9B", size = 0.8) +
  annotate("text", x = 0.07, y = 0.82, label =  "Greater bias when between person differences explain a \nlarger % of total variation in both outcome and predictors.",  hjust = 0) +
  geom_curve(
    aes(x = 0.4, y = 0.85, xend = 0.68, yend = 0.87), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = -0.2
  ) + 
  annotate("text", x = 0.38, y = 0.54, label = "Low bias if between person differences explain a small \n% of total variation in outcome, predictors, or both.",  hjust = 0) +
    geom_curve(
    aes(x = 0.37, y = 0.54, xend = 0.33, yend = 0.51), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = 0.1
  ) +
  # Adding a diagonal line to illustrate the linear relationship
  geom_abline(slope = 0.5, intercept = 0.52, color = "#FE4365",  size = 0.8)

ggsave("Figures/Figure4_Example2.jpg",
  plot = p1,
  width = 10.2,
  height = 4.5)

@@

```{r}
############# Figure 5 #################
########################################
result <- read.csv("simulation_results.csv")

result <- result[result$A != 0 & result$A != 0.8,]

p3 <- ggplot(result, aes(x = icc, y = auc_c_individual - auc_individual)) +
  geom_point(aes(color = ifelse(auc_c_individual - auc_individual > 0, "Above Zero", "Below or Equal to Zero")), 
             alpha = 0.3, size = 0.8, show.legend = FALSE) +
  scale_color_manual(values = c("Above Zero" = "#FC9D9A", "Below or Equal to Zero" = "#83AF9B")) +
  theme_minimal() +
  ylab("Performance Difference \n(centered - uncentered)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome (ICC Outcome)") +
  ggtitle("Within-Person Performance") +
  geom_hline(yintercept = 0) + 
  annotate("text", x = 0, y = 0.28, label = "Centered predictors perform better in most cases (within-person)", hjust = 0) +
  annotate("text", x = 0, y = -0.08, label = "Uncentered predictors perform worse in most cases (within-person)", hjust = 0) +
  ylim(c(-0.45,0.35)) +
  geom_curve(
    aes(x = -0.01, y = 0.01, xend = -0.01, yend = 0.25), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.5, 
    color = "#FC9D9A", 
    curvature = 0
  ) +
    geom_curve(
    aes(x = -0.01, y = 0.01, xend = -0.01, yend = -0.1), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.5, 
    color = "#83AF9B", 
    curvature = 0
  )

p4 <- ggplot(result, aes(x = icc, y = auc_c - auc)) +
  geom_point(aes(color = ifelse(auc_c - auc > 0, "Above Zero", "Below or Equal to Zero")), 
             alpha = 0.3, size = 0.8,show.legend = FALSE) +
  scale_color_manual(values = c("Above Zero" = "#FC9D9A", "Below or Equal to Zero" = "#83AF9B")) +
  theme_minimal() +
  ylab("Performance Difference \n(centered - uncentered)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome (ICC Outcome)") +
  ggtitle("") +
  geom_hline(yintercept = 0) + 
  annotate("text", x = 0, y = 0.28, label = "Centered predictors perform better when between person differences explain a small % of total variance (overall)", hjust = 0) +
  annotate("text", x = 0.01, y = -0.39, label = "Uncentered predictors perform better when between person differences explain a large % of total variance (overall)", hjust = 0) +
  ylim(c(-0.45,0.35)) +
   ggtitle("Overall Performance (Within and Between-Person)")  +
    geom_curve(
    aes(x = -0.01, y = 0.01, xend = -0.01, yend = 0.27), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.5, 
    color = "#FC9D9A", 
    curvature = 0
  ) +
    geom_curve(
    aes(x = 0.8, y = 0.01, xend = 0.8, yend = -0.41), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.5, 
    color = "#83AF9B", 
    curvature = 0
  )

combined_plot <- (p3 / p4) +
  plot_annotation(
    title = "Performance difference: centered vs. not-centered",
  ) 
combined_plot

ggsave(  "Figures/Figure5_Example4.jpg",
  plot = combined_plot,
  width = 10.2,
  height = 5.5)

@@

```{r}
############# Figure 6 #################
########################################
result <- read.csv("simulation_results.csv")
result <- result[result$A != 0 & result$A != 0.8,]
p1 <- ggplot(result, aes(x=auc_c, y=auc_c_individual, color = icc_pred)) +
  geom_point(alpha=0.5, size = 1) +
  #scale_size(range = c(.001, 6), name="") +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#2A363B","#83AF9B","#C8C8A9","#F9CDAD","#FC9D9A","#FE4365")) +
  ylab("AUC within-person centered") +
  xlab("AUC centered") +
  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
  geom_hline(yintercept = 0.5)   + 
annotate("text", x = 0.5, y = 0.9, label = "Linear relationship between overall AUC \nand within-person AUC",  hjust = 0) +
  geom_curve(
    aes(x = 0.63, y = 0.87, xend = 0.75, yend = 0.8), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = -0.15
  ) +
  ggtitle("Uncentered Predictors")

p2 <- ggplot(result, aes(x=auc, y=auc_individual, color = icc_pred)) +
  geom_point(alpha=0.5, size = 1) +
  #scale_size(range = c(.001, 6), name="") +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#2A363B","#83AF9B","#C8C8A9","#F9CDAD","#FC9D9A","#FE4365")) +
  ylab("AUC within-person") +
  xlab("AUC") +
  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
  geom_hline(yintercept = 0.5) + 
annotate("text", x = 0.5, y = 0.9, label = "If between person differences explain a large % of variation in predictors and outcome,\nthe overall AUC becomes high while the within-person AUC remains low.",  hjust = 0) +
  geom_curve(
    aes(x = 0.75, y = 0.82, xend = 0.9, yend = 0.6), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = 0.35
  ) +
  ggtitle("Centered Predictors")

combined_plot <- (p2 / p1) +
  plot_annotation(
    title = "Relationship Overall AUC and within-person AUC (A = 0.05 and A = 0.2)",
  ) 
combined_plot

ggsave( "Figures/Figure6_Example3.jpg",
  plot = combined_plot,
  width = 11.5,
  height = 6.5)
```

## Pitfall 2: Mismatch between chosen cross-validation strategy, employment, and actual observations 
```{r }
############ Table 7 #############
##################################

# Merge the p-values with the original dataset for plotting
data <- features_sample

@@

# Create the sorted plot with conditional background color
p2 <- ggplot(data, aes(x = time, y = y)) +
  # Add background shading for subjects with y = 1
  geom_rect(
    data = data %>% filter(!has_y1),
    aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf),
    fill = "#E5E4E2", alpha = 0.05, inherit.aes = T
  ) +
  geom_point(color = "#FC9D9A", size = 0.5) +                       
  facet_wrap(~subject) +               
  labs(
    x = "Time", 
    y = "Outcome", 
    title = paste("Within-person variability over time: \n88 participants (58.6%) do not experience within-person variability. ")
  ) +
  scale_y_continuous(breaks = c(0, 1)) +  # Ensure only 0 and 1 appear on the y-axis
  theme_minimal() +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
  guides(fill = "none")

@@

ggsave( "Figures/Figure7_Example_var.jpg",
  plot = p2,
  width = 8,
  height = 7)

@@

```{r }
############ Table 5 #############
##################################
sim_1$auc_value_base # row-wise
sim_4$auc_value_base # rolling window

@@

test <- sim_1$ind_base[[1]] %>%
  group_by(subject) %>%
  filter(length(unique(true)) > 1) %>%
  filter(sum(true == 1) > 0, sum(true == 0) > 0) %>%
  summarise(
    auc_val = auc(roc(as.numeric(as.character(true)), as.numeric(as.character(pred)), quiet = TRUE))[1],
    .groups = "drop" # Ungroup after summarizing
  )

@@

>>>>

## Pitfall 1: Within person performance Emprircal Example
```{r }
### Descriptives

# Read the CSV files
final_results_df_2_b <- read.csv("final_results_df_2_b.csv")
final_results_df_3_b <- read.csv("final_results_df_3_b.csv")
final_results_df_4_b <- read.csv("final_results_df_4_b.csv")
final_results_df_5_b <- read.csv("final_results_df_5_b.csv")

# Select specific columns
cols_to_select1 <- c("outcome", "overall_prob_outcome", "sd_outcome",
                  "icc_data")
cols_to_select <- c("overall_prob_outcome", "sd_outcome",
                   "icc_data")

df2_selected <- final_results_df_2_b[, cols_to_select1]
df3_selected <- final_results_df_3_b[, cols_to_select]
df4_selected <- final_results_df_4_b[, cols_to_select]
df5_selected <- final_results_df_5_b[, cols_to_select]

# Rename columns to distinguish between dataframes (optional but recommended)
names(df2_selected) <- paste0(names(df2_selected), "_df2")
names(df3_selected) <- paste0(names(df3_selected), "_df3")
names(df4_selected) <- paste0(names(df4_selected), "_df4")
names(df5_selected) <- paste0(names(df5_selected), "_df5")

# Combine the selected dataframes column-wise
combined_df <- cbind(df2_selected, df3_selected, df4_selected, df5_selected)

# View or write the combined data frame
head(combined_df)

combined_df <- combined_df %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))
write.csv(combined_df, "combined_des_binary.csv", row.names = FALSE)

# Read the CSV files
final_results_df_2 <- read.csv("final_results_df_2.csv")
final_results_df_3 <- read.csv("final_results_df_3.csv")
final_results_df_4 <- read.csv("final_results_df_4.csv")
final_results_df_5 <- read.csv("final_results_df_5.csv")

# Select specific columns
cols_to_select1 <- c("outcome", "overall_prob_outcome", "sd_outcome",
                  "icc_data")
cols_to_select <- c("overall_prob_outcome", "sd_outcome",
                   "icc_data")

df2_selected <- final_results_df_2[, cols_to_select1]
df3_selected <- final_results_df_3[, cols_to_select]
df4_selected <- final_results_df_4[, cols_to_select]
df5_selected <- final_results_df_5[, cols_to_select]

# Rename columns to distinguish between dataframes (optional but recommended)
names(df2_selected) <- paste0(names(df2_selected), "_df2")
names(df3_selected) <- paste0(names(df3_selected), "_df3")
names(df4_selected) <- paste0(names(df4_selected), "_df4")
names(df5_selected) <- paste0(names(df5_selected), "_df5")

# Combine the selected dataframes column-wise
combined_df <- cbind(df2_selected, df3_selected, df4_selected, df5_selected)

# View or write the combined data frame
head(combined_df)

combined_df <- combined_df %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))
write.csv(combined_df, "combined_des_continious.csv", row.names = FALSE)

```{r }
############ Table 4 #############
####### Simulation Study #########

# Basic Parameters from Study
n_features <- 10  # Features
n_samples <- 200  # Timepoints
n_subjects <- 150 # Subjects

# Generating Outcome
overall_prob_outcome <- 0.1  # Overall probability of 1 (for the entire dataset) (e.g., tracking depression 0.146)
sd_outcome <- 0.25 # Controls variability BETWEEN different subject (nneds to be smaller than sqrt(overall_prob_outcome*(1-overall_prob_outcome)))

# Generating Features
A <- 0.05 # Relationship between features and outcome
feature_std <- 0.1 # population level feature generating process
B <- 0.8  # Cross-subject variability ("random effect") (added per participants for all timepoints)
C <- 0.45 # Within-subject variability (added within participant for each timepoint)

@@

set.seed(12361488)
features_sample <- create_data(n_features,n_samples,n_subjects,A,feature_std,B,C,overall_prob_outcome,sd_outcome,0,0)
features_sample <- features_sample[[1]]

@@

```{r }
############ Table 5 #############
############# Binary #############

source("Simulation_UploadData.R")
data = read.csv("esm_clean.csv")
set.seed(1234)

data = data  %>% group_by %>%
  group_by(user_id) %>%
  mutate(day = row_number())

data <- data %>%
  group_by(user_id) %>%
  filter(n_distinct(day) >= 150) %>%
  ungroup()
data = data[data$day <= 150,]

sampled_ids <- sample(unique(data$user_id), size = 20, replace = FALSE)

data <- data %>% 
  filter(user_id %in% sampled_ids)

data$user_id <- sapply(data$user_id, match, unique(unlist(data$user_id)))

id_variable = "user_id"   
time_variable = "day"

colnames(data)[colnames(data) == "user_id"] <- "subject"
colnames(data)[colnames(data) == "day"]     <- "time"
#colnames(data)[colnames(data) == "wantedToFeel_goodAboutMe"]  <- "y"

data <- data %>%
  group_by(subject) %>%
  mutate(y = lag(wantedToFeel_goodAboutMe, n = 1))

data <- data[!is.na(data$y),]

## Run strategies
r_1 <- run_simulation_own(
  features_sample = data,
  cv              = "record-wise",
  n_bootstrap     = 1,
  testsize        = 0.3,
  n_features      =  c("feelingBadToGood", "energy", 
                       "thinkingOverAndOver", "lonely", "selfWorth", "appreciating", "stressed"),
  seed = 1234

)

r_3 <- run_simulation_own(
  features_sample = data,
  cv              = "subject-wise",
  n_bootstrap     = 1,
  testsize        = 0.3,
  n_features      =  c("feelingBadToGood", "energy", 
                       "thinkingOverAndOver", "lonely", "selfWorth", "appreciating", "stressed"),
  seed = 1234

)
r_4 <- run_simulation_slidingwindow_own(
  features_sample = data,
  n_bootstrap     = 1,
  windowsize      = 14,
  n_features      = c("feelingBadToGood", "energy",
                      "thinkingOverAndOver", "lonely", "selfWorth", "appreciating", "stressed"),
  seed = 1234
)

table <- rbind(r_1$overall_summary,r_3$overall_summary, r_4$overall_summary)
table$'AUC Overall' <- c(r_1$auc_value,r_3$auc_value,r_4$auc_value)
colnames(table) <-  c("Mean AUC within-person:", "SD AUC within-person:", "% of AUC > 0.5 within-person:", "N included within-person:", "AUC Overall")
rownames(table) <- c("row-wise","subject-wise","rolling-window")
table
```

```{r }
############ Table 5 #############
############# Continious #########

source("Simulation_continous.R")

data = read.csv("esm_clean.csv")
set.seed(1234)

data = data  %>% group_by %>%
  group_by(user_id) %>%
  mutate(day = row_number())

data <- data %>%
  group_by(user_id) %>%
  filter(n_distinct(day) >= 150) %>%
  ungroup()
data = data[data$day <= 150,]

sampled_ids <- sample(unique(data$user_id), size = 20, replace = FALSE)

data <- data %>% 
  filter(user_id %in% sampled_ids)

data$user_id <- sapply(data$user_id, match, unique(unlist(data$user_id)))

id_variable = "user_id"   
time_variable = "day"

colnames(data)[colnames(data) == "user_id"] <- "subject"
colnames(data)[colnames(data) == "day"]     <- "time"
#colnames(data)[colnames(data) == "wantedToFeel_goodAboutMe"]  <- "y"

data <- data %>%
  group_by(subject) %>%
  mutate(y = lag(lonely, n = 1))

data <- data[!is.na(data$y),]

## Run strategies
r_1 <- run_simulation_own(
  features_sample = data,
  cv              = "record-wise",
  n_bootstrap     = 1,
  testsize        = 0.3,
  n_features      =  c("feelingBadToGood", "energy", 
                       "thinkingOverAndOver", "selfWorth", "appreciating", "stressed"),
  seed = 1234

)

r_3 <- run_simulation_own(
  features_sample = data,
  cv              = "subject-wise",
  n_bootstrap     = 1,
  testsize        = 0.3,
  n_features      =  c("feelingBadToGood", "energy", 
                       "thinkingOverAndOver", "selfWorth", "appreciating", "stressed"),
  seed = 1234

)
r_4 <- run_simulation_slidingwindow_own(
  features_sample = data,
  n_bootstrap     = 1,
  windowsize      = 14,
  n_features      = c("feelingBadToGood", "energy",
                      "thinkingOverAndOver", "selfWorth", "appreciating", "stressed"),
  seed = 1234
)

table <- rbind(
  r_1[c("rsq", "cor", "overall_summary.mean_r2", "overall_summary.mean_cor", "overall_summary.sd")],
  r_3[c("rsq", "cor", "overall_summary.mean_r2", "overall_summary.mean_cor", "overall_summary.sd")],
  r_4[c("rsq", "cor", "overall_summary.mean_r2", "overall_summary.mean_cor", "overall_summary.sd")]
)

table <- round(table[,c(5,1,2,3,4)], 2)
table
```

```{r }
## Try out

# Basic Parameters from Study
n_features <- 10  # Features
n_samples <- 200  # Timepoints
n_subjects <- 15 # Subjects

# Generating Outcome
overall_prob_outcome <- 0.4  # Overall probability of 1 (for the entire dataset) (e.g., tracking depression 0.146)
sd_outcome <- 0.3 # Controls variability BETWEEN different subject (nneds to be smaller than sqrt(overall_prob_outcome*(1-overall_prob_outcome)))

# Generating Features
A <- 0.02 # Relationship between features and outcome
feature_std <- 0.1 # population level feature generating process
B <- 0.8  # Cross-subject variability ("random effect") (added per participants for all timepoints)
C <- 0.2 # Within-subject variability (added within participant for each timepoint)

set.seed(12361488)
features_sample <- create_data(n_features, n_samples, n_subjects, A, feature_std, B, C, overall_prob_outcome, sd_outcome, 0,0)

features_sample <- features_sample[[1]]

sim_1 <- run_simulation(features_sample,"row-wise",1, testsize = 0.3)

table <- rbind(sim_1$overall_summary)
table$'AUC Overall' <- c(sim_1$auc_value)

table

library(dplyr)
library(ggplot2)

pred_true = as.data.frame(sim_1$ind)

pred_true <- pred_true %>%
  arrange(subject, true) %>%
  group_by(subject) %>%
  mutate(day = row_number())

library(pROC)
library(dplyr)

# Compute individual AUC
ind_auc <- pred_true %>%
  group_by(subject) %>%
  filter(length(unique(true)) > 1) %>%
  filter(sum(true == 1) > 0, sum(true == 0) > 0) %>%
  summarise(
    auc_val = pROC::auc(pROC::roc(response = true, predictor = pred, direction = "<",positive = "1", levels = c(0,1)))[1],
    .groups = "drop"
  )

# Merge AUCs back into pred_true
pred_true <- left_join(pred_true, ind_auc, by = "subject")
pred_true$auc_val = round(pred_true$auc_val,2)
pred_true$auc_val[is.na(pred_true$auc_val)] = "-"

pred_true$pred  = ifelse(pred_true$pred  > 0.5, 1, 0)

# Add prediction correctness column
pred_true <- pred_true %>%
  mutate(prediction_correct = ifelse(true == pred, "Correct", "Incorrect")) # CHECK AGAIN IF PREDICTION IS 0/1 !!!

plot = ggplot(pred_true, aes(day, subject, color = factor(true), shape = prediction_correct)) + 
  geom_point(size = 2, alpha = 0.8) +
  geom_text(aes(label = ifelse(!duplicated(subject), paste("AUC: ", auc_val), "")),
            x = -Inf, hjust = 0, size = 3, color = "gray40") +  # AUC on the left side
  scale_color_manual(values = c("0" = "#FFB5A7", "1" = "#40916C")) +
  scale_shape_manual(values = c("Correct" = 15, "Incorrect" = 0)) +
  labs(color = NULL, shape = NULL) +  # Removes legend titles
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    axis.text.y = element_blank(),
    axis.text.x = element_blank()     # Removes x-axis tick labels
  ) +
  xlim(-6, 75) +
  labs(y = paste0("Within-Person"), x = "") +
  ggtitle(paste("AUC evaluated over the whole dataset:", round(table$'AUC Overall', 2), "(Between-Person)")) +
  theme(text = element_text('Futura'),axis.title=element_text(size=14)) 
plot
ggsave("WithinBetween_Example.png", 
       width = 6100, height = 3000, units = "px", dpi = 800,plot = plot)

@@

result <- read.csv("simulation_results_0.csv") # see Run_Simulation.R to investigate how results were created

p1 <- ggplot(result[result$A == 0,], aes(x=icc, y=auc, color = icc_pred)) + # color: icc_pred
  geom_point(alpha=0.4, size = 0.8) +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#1B4332", "#40916C", "#74C69D", "#B5E48C", "#FDE68A", "#FFB5A7")) +
  ylab("Prediction Performance (Overall AUC)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome \n(ICC Outcome)") +
  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
  ggtitle("Bias in Performance Estimates (AUC) When No Relationship Exists: Expected AUC = 0.5") +
  annotate("text", x = 0.07, y = 0.82, label =  "Greater bias when between person differences explain a \nlarger % of total variation in both outcome and predictors.",  hjust = 0) +
  geom_curve(
    aes(x = 0.2, y = 0.86, xend = 0.35, yend = 0.93), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = -0.2
  ) + 
  annotate("text", x = 0.38, y = 0.58, label = "Low bias if between person differences explain a small \n% of total variation in outcome, predictors, or both.",  hjust = 0) +
    geom_curve(
    aes(x = 0.37, y = 0.55, xend = 0.33, yend = 0.51), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = 0.1
  ) 
  # Adding a diagonal line to illustrate the linear relationship
ggsave("Figures/Figure4_Example2.jpg",
  plot = p1,
  width = 10.2,
  height = 4.5)

@@

```{r}
############# Figure 5 BINARY #################
########################################
library(ggplot2)
library(tidyr)
library(dplyr)
library(patchwork)

final_results_df_2_b = read.csv("final_results_df_2_b.csv")
final_results_df_3_b = read.csv("final_results_df_3_b.csv")
final_results_df_4_b = read.csv("final_results_df_4_b.csv")
final_results_df_5_b = read.csv("final_results_df_5_b.csv")

plot_data = function(data_viz, spanpar){
  # Reshape to long format
  df_long <- data_viz %>%
    pivot_longer(
      cols = c(`shuffle_AUC.`, `movingwindowshuffle_AUC.`),
      names_to = "AUC_type",
      values_to = "AUC"
    )

  p = ggplot(df_long, aes(x = icc_data, y = AUC, color = AUC_type)) +
    geom_point(alpha = 0.8, size = 1) +
    geom_smooth(se = FALSE, method = "loess", span = spanpar) +
    theme_minimal() +
    scale_color_manual(
      values = c(
        "shuffle_AUC." = "#40916C",
        "movingwindowshuffle_AUC." = "#4D8FAC"
      ), 
      labels = c(
        "shuffle_AUC." = "No true Relationship (Record-Wise)",
        "movingwindowshuffle_AUC." = "No true Relationship (Moving-Window)"
      )
    ) +
    geom_hline(yintercept = 0.5, color = "grey", size = 0.8) +
    labs(color = "") +
    ylim(0.4, 0.95) +
    ggtitle(paste0("N: ", max(data_viz$num_subjects),
                   ", Timepoints: ", max(data_viz$num_samples),
                   ",\nICC Predictors (mean): ", round(data_viz$icc_pred, 2))) +
    ylab("Overall AUC") +
    xlab("ICC Outcome")

  return(p)
}

spanpar = 0.6

p2 = plot_data(final_results_df_2_b, spanpar)
p3 = plot_data(final_results_df_3_b, spanpar)
p4 = plot_data(final_results_df_4_b, spanpar)
p5 = plot_data(final_results_df_5_b, spanpar)

# Combine all plots and collect legends at the bottom
final_plot <- (p2 + p3 + p4 + p5) + 
  plot_layout(ncol = 2, guides = "collect") & 
  theme(legend.position = "bottom")

# Add title and optional styling
final_plot <- final_plot +
  plot_annotation(
    title = "Predictive Performance Shuffled Data Binary Outcomes",
    theme = theme(
      plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
    )
  )

print(final_plot)
ggsave("Figures/Figure5_NullModelBinary.jpg",
  plot = final_plot,
  width = 7,
  height = 6)
```

```{r}
############# Figure 6 Continiuous #################
########################################
library(ggplot2)
library(tidyr)
library(dplyr)
library(patchwork)

final_results_df_2 = read.csv("final_results_df_2.csv")
final_results_df_3 = read.csv("final_results_df_3.csv")
final_results_df_4 = read.csv("final_results_df_4.csv")
final_results_df_5 = read.csv("final_results_df_5.csv")

plot_data = function(data_viz, spanpar){
  # Reshape to long format
  df_long <- data_viz %>%
    pivot_longer(
      cols = c(shuffle_cor,movingwindowshuffle_cor),
       names_to = "type",
      values_to = "Correlation"
    )

  p = ggplot(df_long, aes(x = icc_data, y = Correlation, color = type)) +
    geom_point(alpha = 0.8, size = 1) +
    geom_smooth(se = FALSE, method = "loess", span = spanpar) +
    theme_minimal() +
    scale_color_manual(values = c(
      "shuffle_cor" = "#40916C",
      "movingwindowshuffle_cor" = "#4D8FAC"
    ), labels = c("shuffle_cor" = "No true Relationship (Record-Wise)", 
                  "movingwindowshuffle_cor" ="No true Relationship (Moving-Window)"
    )) +
    labs(color = "") +
    ggtitle(paste0("N: ", max(data_viz$num_subjects),
                   ", Timepoints: ", max(data_viz$num_samples),
                   ",\nICC Predictors (mean): ", round(data_viz$icc_pred, 2))) +
    ylab("Overall Correlation") +
    xlab("ICC Outcome") +
    ylim(0, 0.7)

  return(p)
}

spanpar = 0.4

p2 = plot_data(final_results_df_2, spanpar)
p3 = plot_data(final_results_df_3, spanpar)
p4 = plot_data(final_results_df_4, spanpar)
p5 = plot_data(final_results_df_5, spanpar)

# Combine the plots and collect the legend
final_plot2 <- (p2 + p3 + p4 + p5) +
  plot_layout(ncol = 2, guides = "collect") & 
  theme(legend.position = "bottom")

# Add title and optional styling
final_plot2 <- final_plot2 +
  plot_annotation(
    title = "Predictive Performance Shuffled Data Continuous Outcomes",
    theme = theme(
      plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
    )
  )

print(final_plot2)
ggsave("Figures/Figure6_NullModelContinious.jpg",
  plot = final_plot2,
  width = 7,
  height = 6)

@@

## Pitfall 2: Mismatch between chosen cross-validation strategy, employment, and actual observations 
```{r }
############ Figure 7 #############
##################################

# Generating Features
A <- 0.05 # Relationship between features and outcome
feature_std <- 0.1 # population level feature generating process
B <- 0.8  # Cross-subject variability ("random effect") (added per participants for all timepoints)
C <- 0.45 # Within-subject variability (added within participant for each timepoint)

set.seed(12361488)
features_sample <- create_data(n_features,90,150,A,feature_std,B,C,overall_prob_outcome,sd_outcome,0,0)

# Merge the p-values with the original dataset for plotting
data <- features_sample[[2]]

@@

# Create the sorted plot with conditional background color
p2 <- ggplot(data, aes(x = time, y = y)) +
  # Add background shading for subjects with y = 1
  geom_rect(
    data = data %>% filter(!has_y1),
    aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf),
    fill = "#E5E4E2", alpha = 0.05, inherit.aes = T
  ) +
  geom_point(color = "#FC9D9A", size = 0.5) +                       
  facet_wrap(~subject) +               
  labs(
    x = "Time", 
    y = "Outcome", 
    title = paste("Within-person variability over time: \n78 participants (52%) do not experience within-person variability. ")
  ) +
  scale_y_continuous(breaks = c(0, 1)) +  # Ensure only 0 and 1 appear on the y-axis
  theme_minimal() +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
  guides(fill = "none")

@@

# ggsave( "Figures/Figure7_Example_var.jpg",
#   plot = p2,
#   width = 8,
#   height = 7)

@@

```{r }
############ Table 7 #############
##################################
sim_1$auc_value_base # row-wise
sim_4$auc_value_base # rolling window

@@

test <- sim_4$ind_base[[1]] %>%
  group_by(subject) %>%
  filter(length(unique(true)) > 1) %>%
  filter(sum(true == 1) > 0, sum(true == 0) > 0) %>%
  summarise(
    auc_val = pROC::auc(pROC::roc(as.numeric(as.character(true)), as.numeric(as.character(pred)), quiet = TRUE))[1],
    .groups = "drop" # Ungroup after summarizing
  )

@@

```{r}
############# Figure 9 BINARY #################
########################################
library(ggplot2)
library(tidyr)
library(dplyr)
library(patchwork)

final_results_df_2_b = read.csv("final_results_df_2_b.csv")
final_results_df_3_b = read.csv("final_results_df_3_b.csv")
final_results_df_4_b = read.csv("final_results_df_4_b.csv")
final_results_df_5_b = read.csv("final_results_df_5_b.csv")

plot_data = function(data_viz,spanpar){
  # Reshape to long format
  df_long <- data_viz %>%
    pivot_longer(
      cols = c(`real_AUC.random.intercept.only.`, `real_AUC.`,`movingwindow_AUC.`,`subjectwise_AUC.`,`movingwindow_AUC.random.intercept.only.`),
      names_to = "AUC_type",
      values_to = "AUC"
    )
  p = ggplot(df_long, aes(x = icc_data, y = AUC, color = AUC_type)) +
    geom_point(alpha = 0.8, size = 1) +
    geom_smooth(se = FALSE, method = "loess",span = spanpar) +
    theme_minimal() +
    scale_color_manual(values = c(
      "real_AUC.random.intercept.only." ="#B5E48C" ,
      "real_AUC." = "#74C69D",
      "movingwindow_AUC." ="#9EC5E7",
      "movingwindow_AUC.random.intercept.only." = "#A2D2FF",
      "subjectwise_AUC." ="#FFB5A7"
    ), 
    labels = c(
               "real_AUC.random.intercept.only."="Random Intercept only (Record-Wise)",
               "real_AUC." ="Random Forest (Record-Wise) ",
               "movingwindow_AUC." ="Random Forest (Moving-Window)",
               "subjectwise_AUC." ="Random Forest (Subject-Wise)",
               "movingwindow_AUC.random.intercept.only." = "Random Intercept only (Moving-Window)")) +
    theme(legend.position = "none") +
    geom_hline(yintercept = 0.5, color = "grey", size = 0.8) +
    labs(color = "Model") +
    ylim(0.4,0.95) +
    ggtitle(paste0("N: ", max(data_viz$num_subjects),", ", "Timepoints: ", max(data_viz$num_samples),", ","\n",  "ICC Predictors (mean): ", round(data_viz$icc_pred,2))) +
    ylab("Overall AUC") +
    xlab("ICC Outcome")
  return(p)
}

spanpar = 0.6

p2 = plot_data(final_results_df_2_b, spanpar)
p3 = plot_data(final_results_df_3_b, spanpar)
p4 = plot_data(final_results_df_4_b, spanpar)
p5 = plot_data(final_results_df_5_b, spanpar)

# Combine all plots and collect legends at the bottom
final_plot <- (p2 + p3 + p4 + p5) + 
  plot_layout(ncol = 2, guides = "collect") & 
  theme(legend.position = "right")

# Add title and optional styling
final_plot <- final_plot +
  plot_annotation(
    title = "Predictive Performance Baseline Comparison Binary Outcomes",
    theme = theme(
      plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
    )
  )

print(final_plot)
ggsave("Figures/Figure9_BaselineBinary.jpg",
  plot = final_plot,
  width = 9,
  height = 8)
```

```{r}
############# Figure 10 Continiuous #################
########################################
library(ggplot2)
library(tidyr)
library(dplyr)
library(patchwork)

final_results_df_2 = read.csv("final_results_df_2.csv")
final_results_df_3 = read.csv("final_results_df_3.csv")
final_results_df_4 = read.csv("final_results_df_4.csv")
final_results_df_5 = read.csv("final_results_df_5.csv")
